---
title: "Rejction Sampling"
author: "Kuncheng Fan & Pin Huang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rejction Sampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Abstract

Rejection sampling is a basic technique used to generate observations from a distribution. It is also commonly called the acceptance-rejection method or “accept-reject algorithm.”
Simply speaking, if the user has a function he or she is trying to sample from, whose functional form is well known, basically accept the sample by generating a uniform random number at any x and accept it if the value if below the value of the function at that x.

The process: 

1. Draw x uniformly from [Xmin, Xmax] 

2. Draw y uniformly from [0, Ymax]

3. if y < f(x), accept the sample

4. otherwise reject it

5. repeat

```{r, include=FALSE}
library(ggplot2)
```


## Single variable Rejection Sampling

This function uses rejection sampling to sample from given one variable pdf. The function has to be an appropriate pdf, otherwise the function will output error. Notice if the input pdf is discrete, this function cannot detect whether the input is a true pdf. The default of `discrete` is FALSE. This function will automatically find the bound of pdf's domain if it is between [-50, 50], otherwise user should input the bound. The defaults of bound `lb` and `ub` are -Inf and Inf. Input bounds for bounded pdf results faster running.

```{r, include = FALSE}
oneDsample <- function(f, N, lb = -Inf, ub = Inf, discrete = FALSE) {
  bdtest <- seq(-50,50,0.0001)
  if (f(-50) == 0 & f(50) == 0 & mean(f(bdtest)) > 0){
    lb = min(bdtest[which(f(bdtest)>0)])
    ub = max(bdtest[which(f(bdtest)>0)])
  }
  if (f(-50) == 0 & f(50) > 0 & mean(f(bdtest)) > 0){
    lb = min(bdtest[which(f(bdtest)>0)])
  }
  if (f(-50) > 0 & f(50) == 0 & mean(f(bdtest)) > 0){
    ub = max(bdtest[which(f(bdtest)>0)])
  }
  if (discrete == TRUE){
    warning("We cannot test whether a discrete function is a pdf")
  }
  if (discrete == FALSE){
    if (abs(integrate(f, -Inf, Inf)$val - 1) > 0.001) {
      stop("Error: not a pdf. The area under the function you given should be 1")
    }
  }
  if (lb != -Inf & ub != Inf){
    maxf <- max(f(runif(100000,lb,ub)))
    ones = c()
    n = 0
    while (n < N) {
      one <-runif(1,lb,ub)
      if (runif(1, 0, maxf) < f(one)){
        ones = c(ones, one)
        n = n + 1
      }
    }
    return(data.frame(x=ones))
  }
  else {
    x <- seq(-5000,5000, 0.001)
    maxf <- max(f(x))
    mu=x[which(f(x)==maxf)]
    sd = 2/maxf
    C = 2*maxf/dnorm(mu,mu,sd)
    ones = c()
    n = 0
    while (n < N) {
      one = rnorm(1, mu, sd)
      if (runif(1, 0, C * dnorm(one,mu,sd)) < f(one)){
        ones = c(ones, one)
        n = n + 1
      }
    }
    return(data.frame(x=ones))
  }
}
```

**Sample from bounded pdf**
```{r}
f <- function(x) {ifelse(0 < x & x < 1, 4*x^3, 0)}
a <- oneDsample(f,10000)
ggplot(a,aes(x)) + geom_density() + stat_function(fun = f, color = "red")
```

**Sample from unbounded pdf**

```{r}
f <- function(x) {1/pi/(1+x^2)}
a <- oneDsample(f,20000)
ggplot(a,aes(x)) + geom_density() + stat_function(fun = f, color = "red")
```

## Conditional Variable Rejection Sampling

In terms of the principle of OneDsample function, TwoDsample function is also based on rejection sampling and designed to sample from (X, Y) given the joint density function. Initially, the user need to input an appropriate joint density function, otherwise the function will output error. Compared with OneDsample, there is no judging criteria for discrete and this function cannot automatically find the bound for this given function. So, the user need to input the lower and upper bounds for both X and Y (if exists). According to the limitation of normal distribution, the defaults of bound are set as -5000 and 5000. 

```{r, include = FALSE}
twoDsample <- function(f, N, lbx=-5000, ubx=5000, lby=-5000, uby=5000) {
  library(MASS)
  library(cubature)
  if (abs(adaptIntegrate(f, c(lbx, lby), c(ubx, uby), maxEval=10000)$integral - 1) > 0.001) {
    stop("Error: Bound is missing/wrong or the function is not a pdf. The area under the function you given should be 1")
  }
  if (lbx != -5000 & ubx != 5000 & lby != -5000 & uby != 5000){
    maxf <- max(replicate(100000,f(c(runif(1,lbx,ubx),runif(1,lby,uby)))))
    twos = c()
    n = 0
    while (n < N) {
      two <- c(runif(1,lbx,ubx),runif(1,lby,uby))
      if (runif(1, 0, maxf) < f(two)){
        twos = c(twos, two)
        n = n+1
      }
    }
    data.frame(x=twos[c(seq(1,length(twos)-1,2))],y=twos[c(seq(2,length(twos),2))])
  }
  else{
    dmvnorm = function(x,mu,sig){
      x1 = x[1]
      x2 = x[2]
      mu1 = mu[1]
      mu2 = mu[2]
      sig1 = sig[1]
      sig2 = sig[2]
      exp(-1/2*((x1-mu1)^2/sig1^2 - 2*(x1-mu1)*(x2-mu2)/sig1/sig2 + (x2-mu2)^2/sig2^2))/(2*pi*sig1*sig2)
    }
    op = optim(c((ubx+lbx)/2,(uby+lby)/2), f, control = list(fnscale = -1))
    maxf = op$value
    mu = c(op$par)
    sd = 2/maxf
    C = maxf/dmvnorm(c(mu[1],mu[2]),c(mu[1],mu[2]),c(sd,sd))
    twos = c()
    n = 0
    while (n < N) {
      two = mvrnorm(1, mu, matrix(c(sd,0,0,sd),2,2))
      if (runif(1, 0, C * dmvnorm(two,mu,c(sd,sd))) < f(two)){
        twos = c(twos, two)
        n = n + 1
      }
    }
    return(data.frame(x=twos[c(seq(1,length(twos)-1,2))],y=twos[c(seq(2,length(twos),2))]))
  }
}
```

**Sample from bounded pdf**

```{r}
jointPFF <- function(x){
  x1 = x[1]
  x2 = x[2]
  ifelse(0<x1 & x1<1 & 0<x2 & x2<1 & 0<x1+x2 & x1+x2<1, 24*x1*x2, 0)}

a <- twoDsample(f = jointPFF, N=10000, lbx = 0, ubx = 1, lby = 0, uby = 1)
ggplot(a, aes(x, y)) +  geom_density_2d()
```

**Sample from unbounded pdf**

```{r}
f <- function(x){
x1 = x[1]
x2 = x[2]
ifelse(x2>0, 1/pi/(1+x1^2) * 0.05*exp(-0.05*x2), 0)}
 
a <- twoDsample(f = f, N=10000)
ggplot(a, aes(x, y)) +  geom_density_2d()
```

## Evalue

Based on the sampling results from OneDsample or TwoDsample functions, we can use Evalue function to compute the expected value of g(X) or g(X, Y).

```{r, include=FALSE}
Evalue <-function(g,rv){
  len = length(rv)
  if (len == 2){
    meanxy = mean(g(rv$x,rv$y))
    return(meanxy)
  }
  else if (len == 1){
    meanx = mean(g(rv))
    return(meanx)
  }
  else {
    stop("Error: invalid variables format")
  }
}
```

**Example of single variable pdf**

```{r}
betaPDF <- function(x) {
 ifelse(0 < x & x < 1, 2*x, 0)}
 x1 = data.frame(oneDsample(f = betaPDF, N=10000, lb = 0, ub = 1))
 g1 = function(x){x^2}
 Evalue(g1,x1)
``` 

**Example of joint pdf**

```{r}
 jointPDF <- function(x){
 x1 = x[1]
 x2 = x[2]
 ifelse(0<x1 & x1 <1 & 0<x2 & x2<1 & 0<x1+x2 & x1+x2<1, 24*x1*x2, 0)}
 xy = data.frame(twoDsample(f = jointPDF, N=10000, lbx=0, ubx=1, lby=0, uby=1))
 g2 = function(x,y){x^2+y}
 Evalue(g2,xy)
```

## Gprob

Based on the sampling results from OneDsample or TwoDsample functions, we can use Gprob function to compute the probabilities related to X or X and Y.

```{r, include=FALSE}
Gprob <-function(cd,data){
  len = length(data)
  if (len == 2){
    meanxy = mean(cd(data$x,data$y))
    return(meanxy)
  }
  else if (len == 1){
    meanx = mean(cd(data))
    return(meanx)
  }
  else {
    stop("Error: invalid data format")
  }
}
```

**Example of single variable pdf**

```{r}
f <- function(x) {
  ifelse(0 < x & x < 1, 4*x^3, 0)}

x1 = oneDsample(f, 10000)
c1 = function(x){x<0.4}
Gprob(c1,x1)
```

**Example of joint pdf**

```{r}
 f <- function(x){
 x1 = x[1]
 x2 = x[2]
 ifelse(x2>0, 1/pi/(1+x1^2) * 0.05*exp(-0.05*x2), 0)}
 x2 = twoDsample(f = f, N=10000)
 c2 = function(x,y){
  x<0.5 & y<0.2}
 Gprob(c2,x2)
```




